Measuring the Price of Anarchy in Critical Care Unit Interactions

### 2016-07-04 Paper/paper.tex
Modifications to paper.

@@

The effect of state dependent service rates in healthcare has been well studied in isolated hospitals.
In \cite{batt2012doctors} an empirical study is undertaken and service rate slow down is identified.
Furthermore, it is shown that modelling whilst ignoring these state dependent rates leads to errors.
This effect is further identified in \cite{chan2014use, kc2012econometric} where for example the negative effect on patient outcome is revealed.
In \cite{kim2013icu, shmueli2003optimizing} this is expanded to consider a variety of admission policies in two CCUs, in particular the effect of rejection (due to too high occupancy) is measured.
These policies are studied in the setting of a single hospital, and thus, from a game theoretic \cite{Maschler2013} point of view correspond to rational behaviour.
In practice a simplification of rational strategies is managed by policy makers by setting utilisation targets that ensure that hospitals do not run at a level likely to have a high rejection rate (example of these can be seen in \cite{Bevan2006, NHS2013}).

@@

A Critical Care Unit (CCU), also sometimes referred to as an Intensive Therapy Unit or Intensive Therapy Department, is a special ward that is found in most acute hospitals.
It provides intensive care (treatment and monitoring) for people who are critically ill or are in an unstable condition.
CCUs face major challenges, on average, 8\% of patients are refused admission to a CCU because the Unit is full \cite{Report}.
The CCU occupancy rates for some hospitals are reportedly very high
\cite{Mitchell1995,Smith1995} and a shortage of beds has been identified
throughout the United Kingdom.

Many previous researchers have developed queueing models to help manage bed capacities in hospitals \cite{Cooper1974,Dumas1984,Gallivan2011,Gorunescu2002a,Griffiths2012,Harper2002a}.
Also, a vast amount of literature has been devoted to the simulation of CCUs;  \cite{Cahill1999a,Costa2003,Griffiths2004a,Kim1999,Litvak2008,Shahani2008}.

This paper describes part of a project undertaken with managers from the Aneurin Bevan University Health Board (ABUHB), which is an NHS Wales organization in South Wales, that serves 21\% of the total population of Wales \cite{Board}.
Critical care is delivered on two sites, at the Nevill Hall hospital in Abergavenny and at the Royal Gwent hospital in Newport.
For the remainder of this paper, the Nevill Hall hospital will be referred to as NH and the Royal Gwent hospital as RG.

The main proposition of the work requested by the ABUHB was to develop a mathematical model of bed occupancies at the CCUs at RG and NH.
After an initial analysis of the data, behavioural aspects became apparent; for example, delaying patients discharge if there was no pressure on CCU beds, or admitting fewer patients if bed occupancy levels were high.
As a result of this, a state-dependent queueing model has been developed, which includes the dependency of admission rate on actual occupancy \cite{williams2015mathematical}.
This state dependent model was applied to both NH and RG separately.
It is however obvious that the actions of one CCU impact on the other CCU, as diversion of patients from one CCU to the other sometimes occurs.
A pictorial representation of the situation is given in Figure \ref{diagramofdiversion}.

@@

The work presented in this paper contributes to the growing body of literature by applying state dependent queueing models in a game theoretical context to CCU interaction.
In particular this consideration allows for the investigation of targets imposed by central control \cite{Bevan2006}.
The findings of this work justify and identify a choice of targets that align the interests of the individual hospitals with social welfare.

@@

\begin{figure}[!htbp]
\begin{center}
    \begin{tikzpicture}[scale=.7, every node/.style={scale=0.7}]
        % -----------------------------------------------
        % Boundaries ------------------------------------
        % -----------------------------------------------
        \draw [dashed, red, line width=1mm, opacity=.3] (12,2) -- (12,-20);
        \draw [dashed, red, line width=1mm, opacity=.3] (-2,-12) -- (20,-12);
        %\tikzstyle{state}=[ellipse, draw, fill=blue!10, minimum width=2cm];
        %\tikzstyle{state}=[rectangle, draw, fill=blue!10, minimum width=2cm];
        \tikzstyle{state}=[minimum width=2cm, font=\boldmath];
        %\draw  (0,0) rectangle (4,3);
        %\draw [dashed, red, thick] (0,1) -- (4,1);
        %\draw [dashed, red, thick] (3,0) -- (3,3);
        % -----------------------------------------------
        % First row--------------------------------------
        % -----------------------------------------------
        \node (aa) [state] at (0,0) {$(0,0)$};
        \node (ab) [state] at ($(aa) + (3,0)$) {$(0,1)$};
        \node (ac) [state] at ($(ab) + (3,0)$) {$(0,2)$};
        \node (ellipsisa1) at ($(ac) + (3,0)$) {$\dots$};
        \node(ak) [state] at ($(ellipsisa1) + (3,0)$) {$(0,K_{\RG})$};
        \node (ellipsisa2) at ($(ak) + (3,0)$) {$\dots$};
        \node(az) [state] at ($(ellipsisa2) + (3,0)$) {$(0,c_{\RG})$};
        % Transitions -----------------------------------
        % On Row:
        \draw (aa) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(l,l)}$} (ab);
        \draw (aa) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (ab);

@@

The stochastic transition rate matrix $Q=Q(c_{NH},c_{RG})$ of the continuous-time Markov chain \cite{Stewart2009} has entries $q_{ij}=q_{(u_i,v_i),(u_j,v_j)}$ which is the rate at which a transition from state $i$ to state $j$ occurs. The transition rates are given by (\ref{transitions}) and are illustrated diagrammatically in Figure \ref{mc}.

@@

Utilities will be of interest when this queueing theoretical model will be inserted in the game theoretical model.
Throughput of patients is a natural choice of utility given that most hospitals are financially rewarded per served patient \cite{Pate2009}.
For each threshold pair $(K_{\NH},K_{\RG})$, the utilisation rate $U_H$ and throughput $T_H$ can easily be obtained for each CCU: $H\in\{\text{NH},\text{RG}\}$, using the following formulas:

@@

\begin{figure}[!htbp]
$$
\begin{array}{cc}
\includegraphics[width=.5\textwidth]{./Images/NH.pdf}&\includegraphics[width=.5\textwidth]{./Images/RG.pdf}\\
H=\NH&H=\RG
\end{array}
$$
\caption{Steady state probabilities for $H\in\{\NH,\RG\}$ with $(K_{\NH}, K_{\RG})=(6,12)$} \label{steady_state_probs}
\end{figure}

@@

\begin{figure}[!htbp]
$$
\begin{array}{cc}
\includegraphics[width=.5\textwidth]{./Images/NH_1_12.pdf}&\includegraphics[width=.5\textwidth]{./Images/RG_1_12.pdf}\\
H=\NH&H=\RG
\end{array}
$$
\caption{Steady state probabilities for $H\in\{\NH,\RG\}$ with $(K_{\NH}, K_{\RG})=(1,12)$} \label{steady_state_probs_1_12}
\end{figure}

@@

This game is equivalent to a bi matrix game with restriction to pure strategies where both players aim to get their utilisation as close as possible to a certain target.
As such a Nash Equilibrium is not guaranteed by traditional game theoretical results \cite{Nash1950}, which guarantee the existence of equilibria in mixed strategies.
Based on discussions with ABUHB, long term threshold policies are a realistic consideration and so a pure strategy space is used.

@@

The function $f_H$ is well defined as it maximises a continuous function over a
finite discrete set.  In case of multiple values that minimize $U_H$, it is
assumed that $f_H$ returns the lowest such value, this is consistent with the
Price of Anarchy (PoA) being a theoretical upper bound of the effect of
uncoordinated behaviour \cite{TimRoughgarden}.

@@

\begin{figure}[!htbp]
    \begin{center}
        \includegraphics[width=.7\textwidth]{./Images/best_responses_ex_for_proof.pdf}
    \end{center}
    \caption{Plots of $f_{\NH}(K_{\RG})$ (crosses) and $f_{\RG}(K_{\NH})$ (diamonds)}\label{exampleforproof}
\end{figure}

@@

\begin{figure}[!hbtp]
    \begin{center}
        \begin{tikzpicture}[scale=.8, every node/.style={scale=0.8}]
            \coordinate (origin) at (0,0);
                % Coordinate system
                \draw [very thick, ->] (origin) -- ($(origin) + (0,5)$) node[anchor=south east] {\Large \(\bar\lambda_{K_{\RG}}\)};
                \draw [very thick, ->] (origin) -- ($(origin) + (5,0)$) node[anchor=west] {\Large Occupancy at RG};
                % lambda
                \draw ($(origin) + (0,2)$) -- ($(origin) + (2,2)$) -- ($(origin) + (2,4)$) -- ($(origin) + (5,4)$);
                \node at ($(origin) + (1,2.5)$) {\(\lambda_{\NH}^{(h,l)}\)};
                \node at ($(origin) + (3.5,4.5)$) {\(\lambda_{\NH}^{(h,h)}\)};
                % Effect of KRG
                \draw [dashed] ($(origin) + (0,2)$) -- ($(origin) + (4,2)$) -- ($(origin) + (4,4)$);
                \draw [dashed] ($(origin) + (2,0)$) -- ($(origin) + (2,2)$);
                \draw [dashed] ($(origin) + (4,0)$) -- ($(origin) + (4,2)$);
                \draw [very thick, ->] ($(origin) + (2.5,3)$) -- ($(origin) + (3.5,3)$);
                \node at ($(origin) + (2,0)$) [below] {\(K_{\RG}\)};
                \node at ($(origin) + (4,0)$) [below] {\(K_{\RG}+1\)};

            \coordinate (origin) at (9,0);
                % Coordinate system
                \draw [very thick, ->] (origin) -- ($(origin) + (0,5)$) node[anchor=south east] {\Large \(\bar\lambda_{K_{\RG}}\)};
                \draw [very thick, ->] (origin) -- ($(origin) + (5,0)$) node[anchor=west] {\Large Occupancy at RG};
                % lambda
                \draw ($(origin) + (0,2)$) -- ($(origin) + (2,2)$) -- ($(origin) + (2,4)$) -- ($(origin) + (5,4)$);
                \node at ($(origin) + (1,2.5)$) {\(\lambda_{\NH}^{(l,l)}\)};
                \node at ($(origin) + (3.5,4.5)$) {\(\lambda_{\NH}^{(l,h)}\)};
                % Effect of KRG
                \draw [dashed] ($(origin) + (0,2)$) -- ($(origin) + (4,2)$) -- ($(origin) + (4,4)$);
                \draw [dashed] ($(origin) + (2,0)$) -- ($(origin) + (2,2)$);
                \draw [dashed] ($(origin) + (4,0)$) -- ($(origin) + (4,2)$);
                \draw [very thick, ->] ($(origin) + (2.5,3)$) -- ($(origin) + (3.5,3)$);
                \node at ($(origin) + (2,0)$) [below] {\(K_{\RG}\)};
                \node at ($(origin) + (4,0)$) [below] {\(K_{\RG}+1\)};
        \end{tikzpicture}
    \end{center}
    \caption{The effect of \(K_{\RG}\) on \(\bar \lambda\).}\label{fig:effect_of_KRG_on_arrival_rate}
\end{figure}

@@

\begin{figure}[!hbtp]
    \begin{center}
        \begin{tikzpicture}[scale=.8, every node/.style={scale=0.8}]
            \coordinate (origin) at (0,0);
                % Coordinate system
                \draw [very thick, ->] (origin) -- ($(origin) + (0,5)$) node[anchor=south east] {\Large \(\bar\lambda_{K_{\NH}}\)};
                \draw [very thick, ->] (origin) -- ($(origin) + (5,0)$) node[anchor=west] {\Large Occupancy at NH};
                % lambda
                \draw ($(origin) + (0,4)$) -- ($(origin) + (2,4)$) -- ($(origin) + (2,2)$) -- ($(origin) + (5,2)$);
                \node at ($(origin) + (1,3.5)$) {\(\lambda_{\NH}^{(h,l)}\)};
                \node at ($(origin) + (3.5,1.5)$) {\(\lambda_{\NH}^{(h,h)}\)};
                % Effect of KNH
                \draw [dashed] ($(origin) + (0,4)$) -- ($(origin) + (4,4)$) -- ($(origin) + (4,2)$);
                \draw [dashed] ($(origin) + (2,0)$) -- ($(origin) + (2,4)$);
                \draw [dashed] ($(origin) + (4,0)$) -- ($(origin) + (4,4)$);
                \draw [very thick, ->] ($(origin) + (2.5,3)$) -- ($(origin) + (3.5,3)$);
                \node at ($(origin) + (2,0)$) [below] {\(K_{\NH}\)};
                \node at ($(origin) + (4,0)$) [below] {\(K_{\NH}+1\)};

            \coordinate (origin) at (9,0);
                % Coordinate system
                \draw [very thick, ->] (origin) -- ($(origin) + (0,5)$) node[anchor=south east] {\Large \(\bar\lambda_{K_{\NH}}\)};
                \draw [very thick, ->] (origin) -- ($(origin) + (5,0)$) node[anchor=west] {\Large Occupancy at NH};
                % lambda
                \draw ($(origin) + (0,4)$) -- ($(origin) + (2,4)$) -- ($(origin) + (2,2)$) -- ($(origin) + (5,2)$);
                \node at ($(origin) + (1,3.5)$) {\(\lambda_{\NH}^{(l,l)}\)};
                \node at ($(origin) + (3.5,1.5)$) {\(\lambda_{\NH}^{(l,h)}\)};
                % Effect of KNH
                \draw [dashed] ($(origin) + (0,4)$) -- ($(origin) + (4,4)$) -- ($(origin) + (4,2)$);
                \draw [dashed] ($(origin) + (2,0)$) -- ($(origin) + (2,4)$);
                \draw [dashed] ($(origin) + (4,0)$) -- ($(origin) + (4,4)$);
                \draw [very thick, ->] ($(origin) + (2.5,3)$) -- ($(origin) + (3.5,3)$);
                \node at ($(origin) + (2,0)$) [below] {\(K_{\NH}\)};
                \node at ($(origin) + (4,0)$) [below] {\(K_{\NH}+1\)};
        \end{tikzpicture}
    \end{center}
    \caption{The effect of \(K_{\NH}\) on \(\bar \lambda\).}\label{fig:effect_of_KNH_on_arrival_rate}
\end{figure}

@@

            % Horizontal inequalities
            \foreach \x in {1.5,2.5,3.5,4.5,5.5}
                {
                \foreach \y in {1,2,3,4,5}
                    {
                        \node at (\x, \y) [red!100] {\textbf{\(\leq\)}};
                    };
                };
        \end{tikzpicture}
    \end{center}
    \caption{Values of \(U_{\NH}\) and the corresponding inequalities for varying cutoff thresholds.}\label{fig:variation_of_UNH}
\end{figure}

@@

Recalling equations (\ref{equ:UNH_as_a_function_of_KRG}-\ref{equ:UNH_as_a_function_of_KNH}) and Figure \ref{fig:variation_of_UNH} have:

@@

The measure used to quantify inefficiency is the Price of Anarchy (PoA) \cite{Koutsoupias1999,TimRoughgarden}, which is the ratio of the social optimum welfare to the welfare of the worst Nash Equilibrium.
That is, the ratio of the largest social welfare, $T^*$ to the smallest social welfare, $\widetilde{T}$, achieved at any Nash Equilibrium. Thus:

@@

The game theoretic model of (\ref{model1}) is solved using exhaustive consideration of best responses whilst taking advantage of the structure identified by the Lemma of Section \ref{queueingandgamemodels}. For any given pair of threshold strategies the matrix equation $\pi Q=0$ is solved by obtaining a basis for the Kernel of $Q$. For the purpose of this paper this is implemented in \cite{sage}.

@@

This model assumes that if the bed occupancy level at both CCUs exceeds a
predetermined threshold, then the admission to the CCU is cancelled. This
cancellation could correspond to sending the patient to a completely different
CCU (outside of the model), moving one of the current CCU patients (ready to be
discharged) to another ward and/or using the post-anesthesia care unit
as a temporary overflow measure. This models corresponds to the first of those
possibilities: the patient is lost (from the point of view of this model).

@@

We see that an extremely large PoA is obtained for $t<0.2$.
For values of $t>0.5$ the PoA is still high: a PoA of 2 corresponds to 100\% less throughput of patients.
These findings seem to give some backing to the targets implemented throughout the NHS \cite{Bevan2006}.

@@

In this work a generic game theoretical model has been presented that accounts for the rational actions of two CCUs.
This game theoretic model is underpinned by a queueing model that takes into account the stochastic nature of queueing systems.
This work extends the application of game theoretic models already present in the literature to healthcare \cite{li2002cooperative, xie2006note}.

@@

\begin{itemize}
    \item The assumptions as to the strategy space of our players is
        restrictive: a single threshold policy might not be optimal (although it
        is present in various pieces of literature on optimal control of
        queueing systems: \cite{naor1969regulation, shone2013comparisons}).
        Indeed, in reality critical care managers could have far more complex
        boundaries for their heuristic decision making;
    \item This model only assumes the presence of two players however in reality
        the system has a variety of stakeholders. Multi player systems could be
        worth considering. This would reflect health boards/hospitals in a
        concentrated area so that interactions are not just between two
        hospitals but between many.
    \item The restriction to pure strategies is influenced by discussions with
        ABUHB and also does not detract from the results presented thanks to the
        Theorem and Lemma of Section \ref{queueingandgamemodels}.  However,
        allowing for mixed strategies could also be of interest, corresponding
        to decision making that is not constant over time. Managers could
        alternate between a variety of behaviours: at time accepting patient
        despite being busy and at other times not;
    \item Patient length of stay is assumed to be dependent on the CCU at which
        they receive service. A further extension of the work would be to use
        the service rate from original CCU (prior to diversion). This would
        require a Markov chain with a higher dimensional state space. In
        practice this corresponds to patient morbidity corresponding to the
        original CCU.
\end{itemize}

@@

The code used in this work can be found here: \url{https://gist.github.com/anonymous/81effc06eea70a9e4e2f}.
The graphics for this paper were obtained using \cite{Hunter:2007, sage} a worksheet with the data and code used for the plots can be found here: \url{https://cloud.sagemath.com/projects/c293aefd-1fdf-4b9c-95f4-75bb77035e42/files/MeasuringThePriceOfAnarchyInCCUInteractions.sagews}.

\newpage
\bibliographystyle{plain}
\bibliography{library}

>>>>

\usepackage{natbib}
\bibliographystyle{agsm}

@@

The effect of state dependent service rates in healthcare has been well studied in isolated hospitals.
In \cite{batt2012doctors} an empirical study is undertaken and service rate slow down is identified.
Furthermore, it is shown that modelling whilst ignoring these state dependent rates leads to errors.
This effect is further identified in \cite{chan2014use, kc2012econometric} where for example the negative effect on patient outcome is revealed.
In \cite{kim2013icu, shmueli2003optimizing} this is expanded to consider a variety of admission policies in two CCUs, in particular the effect of rejection (due to too high occupancy) is measured.
These policies are studied in the setting of a single hospital, and thus, from a game theoretic point of view correspond to rational behaviour (a good game theory reference text is \cite{Maschler2013}).
In practice a simplification of rational strategies is managed by policy makers by setting utilisation targets that ensure that hospitals do not run at a level likely to have a high rejection rate (example of these can be seen in \cite{Bevan2006, NHS2013}).

@@

A Critical Care Unit (CCU), also sometimes referred to as an Intensive Therapy Unit or Intensive Therapy Department, is a special ward that is found in most acute hospitals.
It provides intensive care (treatment and monitoring) for people who are critically ill or are in an unstable condition.
CCUs face major challenges, on average, 8\% of patients are refused admission to a CCU because the Unit is full (\cite{Report}).
The CCU occupancy rates for some hospitals are reportedly very high
(\cite{Mitchell1995,Smith1995}) and a shortage of beds has been identified
throughout the United Kingdom.

Many previous researchers have developed queueing models to help manage bed capacities in hospitals: \cite{Cooper1974,Dumas1984,Gallivan2011,Gorunescu2002a,Griffiths2012,Harper2002a}.
Also, a vast amount of literature has been devoted to the simulation of CCUs;  \cite{Cahill1999a,Costa2003,Griffiths2004a,Kim1999,Litvak2008,Shahani2008}.

This paper describes part of a project undertaken with managers from the Aneurin Bevan University Health Board (ABUHB), which is an NHS Wales organization in South Wales, that serves 21\% of the total population of Wales (\cite{Boarda}).
Critical care is delivered on two sites, at the Nevill Hall hospital in Abergavenny and at the Royal Gwent hospital in Newport.
For the remainder of this paper, the Nevill Hall hospital will be referred to as NH and the Royal Gwent hospital as RG.

The main proposition of the work requested by the ABUHB was to develop a mathematical model of bed occupancies at the CCUs at RG and NH.
After an initial analysis of the data, behavioural aspects became apparent; for example, delaying patients discharge if there was no pressure on CCU beds, or admitting fewer patients if bed occupancy levels were high.
As a result of this, a state-dependent queueing model has been developed, which includes the dependency of admission rate on actual occupancy (\cite{williams2015mathematical}).
This state dependent model was applied to both NH and RG separately.
It is however obvious that the actions of one CCU impact on the other CCU, as diversion of patients from one CCU to the other sometimes occurs.
A pictorial representation of the situation is given in Figure \ref{diagramofdiversion}.

@@

The work presented in this paper contributes to the growing body of literature by applying state dependent queueing models in a game theoretical context to CCU interaction.
In particular this consideration allows for the investigation of targets imposed by central control (\cite{Bevan2006}).
The findings of this work justify and identify a choice of targets that align the interests of the individual hospitals with social welfare.

@@

\begin{figure}[!htbp]
\begin{center}
    \begin{tikzpicture}[scale=.8, every node/.style={scale=0.8}]
        % -----------------------------------------------
        % Boundaries ------------------------------------
        % -----------------------------------------------
        \draw [dashed, red, line width=1mm, opacity=.3] (12,2) -- (12,-20);
        \draw [dashed, red, line width=1mm, opacity=.3] (-2,-12) -- (20,-12);
        %\tikzstyle{state}=[ellipse, draw, fill=blue!10, minimum width=2cm];
        %\tikzstyle{state}=[rectangle, draw, fill=blue!10, minimum width=2cm];
        \tikzstyle{state}=[minimum width=2cm, font=\boldmath];
        %\draw  (0,0) rectangle (4,3);
        %\draw [dashed, red, thick] (0,1) -- (4,1);
        %\draw [dashed, red, thick] (3,0) -- (3,3);
        % -----------------------------------------------
        % First row--------------------------------------
        % -----------------------------------------------
        \node (aa) [state] at (0,0) {$(0,0)$};
        \node (ab) [state] at ($(aa) + (3,0)$) {$(0,1)$};
        \node (ac) [state] at ($(ab) + (3,0)$) {$(0,2)$};
        \node (ellipsisa1) at ($(ac) + (3,0)$) {$\dots$};
        \node(ak) [state] at ($(ellipsisa1) + (3,0)$) {$(0,K_{\RG})$};
        \node (ellipsisa2) at ($(ak) + (3,0)$) {$\dots$};
        \node(az) [state] at ($(ellipsisa2) + (3,0)$) {$(0,c_{\RG})$};
        % Transitions -----------------------------------
        % On Row:
        \draw (aa) edge[out=45,in=135,->,thick] node [above] {\tiny$\lambda_{\RG}^{(l,l)}$} (ab);
        \draw (aa) edge[out=-45,in=-135,<-,thick] node [below] {\tiny$\mu_{\RG}$} (ab);

@@

The stochastic transition rate matrix $Q=Q(c_{NH},c_{RG})$ of the continuous-time Markov chain (\cite{Stewart2009}) has entries $q_{ij}=q_{(u_i,v_i),(u_j,v_j)}$ which is the rate at which a transition from state $i$ to state $j$ occurs. The transition rates are given by (\ref{transitions}) and are illustrated diagrammatically in Figure \ref{mc}.

@@

Utilities will be of interest when this queueing theoretical model will be inserted in the game theoretical model.
Throughput of patients is a natural choice of utility given that most hospitals are financially rewarded per served patient (\cite{Pate2009}).
For each threshold pair $(K_{\NH},K_{\RG})$, the utilisation rate $U_H$ and throughput $T_H$ can easily be obtained for each CCU: $H\in\{\text{NH},\text{RG}\}$, using the following formulas:

@@

\begin{figure}[!htbp]
$$
\begin{array}{cc}
\includegraphics[width=.5\textwidth]{./Images/NH.pdf}&\includegraphics[width=.5\textwidth]{./Images/RG.pdf}\\
H=\NH&H=\RG
\end{array}
$$
\caption{Steady state probabilities for each hospital ($H\in\{\NH,\RG\}$) with different threshold pairs ($(K_{\NH}, K_{\RG})=(6,12)$)} \label{steady_state_probs}
\end{figure}

@@

\begin{figure}[!htbp]
$$
\begin{array}{cc}
\includegraphics[width=.5\textwidth]{./Images/NH_1_12.pdf}&\includegraphics[width=.5\textwidth]{./Images/RG_1_12.pdf}\\
H=\NH&H=\RG
\end{array}
$$
\caption{Steady state probabilities for each hospital ($H\in\{\NH,\RG\}$) with different threshold pairs ($(K_{\NH}, K_{\RG})=(1,12)$)} \label{steady_state_probs_1_12}
\end{figure}

@@

This game is equivalent to a bi matrix game with restriction to pure strategies where both players aim to get their utilisation as close as possible to a certain target.
As such a Nash Equilibrium is not guaranteed by traditional game theoretical results (\cite{Nash1950}), which guarantee the existence of equilibria in mixed strategies.
Based on discussions with ABUHB, long term threshold policies are a realistic consideration and so a pure strategy space is used.

@@

The function $f_H$ is well defined as it maximises a continuous function over a
finite discrete set.  In case of multiple values that minimize $U_H$, it is
assumed that $f_H$ returns the lowest such value, this is consistent with the
Price of Anarchy (PoA) being a theoretical upper bound of the effect of
uncoordinated behaviour (\cite{TimRoughgarden}).

@@

\begin{figure}[!htbp]
    \begin{center}
        \includegraphics[width=.7\textwidth]{./Images/best_responses_ex_for_proof.pdf}
    \end{center}
    \caption{Plots of best response functions: $f_{\NH}(K_{\RG})$ (crosses) and $f_{\RG}(K_{\NH})$ (diamonds)}\label{exampleforproof}
\end{figure}

@@

\begin{figure}[!hbtp]
    \begin{center}
        \begin{tikzpicture}[scale=.8, every node/.style={scale=0.8}]
            \coordinate (origin) at (0,0);
                % Coordinate system
                \draw [very thick, ->] (origin) -- ($(origin) + (0,5)$) node[anchor=south east] {\Large \(\bar\lambda_{K_{\RG}}\)};
                \draw [very thick, ->] (origin) -- ($(origin) + (5,0)$) node[anchor=west] {\Large Occupancy at RG};
                % lambda
                \draw ($(origin) + (0,2)$) -- ($(origin) + (2,2)$) -- ($(origin) + (2,4)$) -- ($(origin) + (5,4)$);
                \node at ($(origin) + (1,2.5)$) {\(\lambda_{\NH}^{(h,l)}\)};
                \node at ($(origin) + (3.5,4.5)$) {\(\lambda_{\NH}^{(h,h)}\)};
                % Effect of KRG
                \draw [dashed] ($(origin) + (0,2)$) -- ($(origin) + (4,2)$) -- ($(origin) + (4,4)$);
                \draw [dashed] ($(origin) + (2,0)$) -- ($(origin) + (2,2)$);
                \draw [dashed] ($(origin) + (4,0)$) -- ($(origin) + (4,2)$);
                \draw [very thick, ->] ($(origin) + (2.5,3)$) -- ($(origin) + (3.5,3)$);
                \node at ($(origin) + (2,0)$) [below] {\(K_{\RG}\)};
                \node at ($(origin) + (4,0)$) [below] {\(K_{\RG}+1\)};
        \end{tikzpicture}
    \end{center}
    \caption{The effect of the threshold (\(K_{\RG}\)) on the effective arrival rate (\(\bar \lambda\)). The same behaviour is exhibited for \(\lambda_{\NH}^{(l,l)}\) and \(\lambda_{\NH}^{(l, h)}\).}\label{fig:effect_of_KRG_on_arrival_rate}
\end{figure}

@@

\begin{figure}[!hbtp]
    \begin{center}
        \begin{tikzpicture}[scale=.8, every node/.style={scale=0.8}]
            \coordinate (origin) at (0,0);
                % Coordinate system
                \draw [very thick, ->] (origin) -- ($(origin) + (0,5)$) node[anchor=south east] {\Large \(\bar\lambda_{K_{\NH}}\)};
                \draw [very thick, ->] (origin) -- ($(origin) + (5,0)$) node[anchor=west] {\Large Occupancy at NH};
                % lambda
                \draw ($(origin) + (0,4)$) -- ($(origin) + (2,4)$) -- ($(origin) + (2,2)$) -- ($(origin) + (5,2)$);
                \node at ($(origin) + (1,3.5)$) {\(\lambda_{\NH}^{(h,l)}\)};
                \node at ($(origin) + (3.5,1.5)$) {\(\lambda_{\NH}^{(h,h)}\)};
                % Effect of KNH
                \draw [dashed] ($(origin) + (0,4)$) -- ($(origin) + (4,4)$) -- ($(origin) + (4,2)$);
                \draw [dashed] ($(origin) + (2,0)$) -- ($(origin) + (2,4)$);
                \draw [dashed] ($(origin) + (4,0)$) -- ($(origin) + (4,4)$);
                \draw [very thick, ->] ($(origin) + (2.5,3)$) -- ($(origin) + (3.5,3)$);
                \node at ($(origin) + (2,0)$) [below] {\(K_{\NH}\)};
                \node at ($(origin) + (4,0)$) [below] {\(K_{\NH}+1\)};
        \end{tikzpicture}
    \end{center}
    \caption{The effect of the threshold (\(K_{\NH}\)) on the effective arrival rate (\(\bar \lambda\)). The same behaviour is exhibited for \(\lambda_{\NH}^{(l,l)}\) and \(\lambda_{\NH}^{(l, h)}\).}\label{fig:effect_of_KNH_on_arrival_rate}
\end{figure}

@@

            % Horizontal inequalities
            \foreach \x in {1.5,2.5,3.5,4.5,5.5}
                {
                \foreach \y in {1,2,3,4,5}
                    {
                        \node at (\x, \y) [red!100] {\textbf{\(\leq\)}};
                    };
                };
        \end{tikzpicture}
    \end{center}
    \caption{Values of utilisation (\(U_{\NH}\)) and the corresponding inequalities for varying cutoff thresholds.}\label{fig:variation_of_UNH}
\end{figure}

@@

Recalling equations
(\ref{equ:UNH_as_a_function_of_KRG}-\ref{equ:UNH_as_a_function_of_KNH}) and
Figure~\ref{fig:variation_of_UNH} gives:

@@

The measure used to quantify inefficiency is the Price of Anarchy (PoA) (\cite{Koutsoupias1999,TimRoughgarden}), which is the ratio of the social optimum welfare to the welfare of the worst Nash Equilibrium.
That is, the ratio of the largest social welfare, $T^*$ to the smallest social welfare, $\widetilde{T}$, achieved at any Nash Equilibrium. Thus:

@@

The game theoretic model of (\ref{model1}) is solved using exhaustive
consideration of best responses whilst taking advantage of the structure
identified by the Lemma of Section~\ref{queueingandgamemodels}. For any given
pair of threshold strategies the matrix equation $\pi Q=0$ is solved by
obtaining a basis for the kernel of $Q$. For the purpose of this paper this is
implemented in Sagemath (\cite{sage}).

@@

This model assumes that if the bed occupancy level at both CCUs exceeds a
predetermined threshold, then the admission to the CCU is cancelled. This
cancellation could correspond to sending the patient to a completely different
CCU (outside of the model), moving one of the current CCU patients (ready to be
discharged) to another ward and/or using the post-anesthesia care unit
as a temporary overflow measure. This models corresponds to the first
possibility: the patient is lost (from the point of view of this model).

@@

We see that an extremely large PoA is obtained for $t<0.2$.
For values of $t>0.5$ the PoA is still high: a PoA of 2 corresponds to 100\% less throughput of patients.
These findings seem to give some backing to the targets implemented throughout the NHS (\cite{Bevan2006}).

@@

In this work a generic game theoretical model has been presented that accounts for the rational actions of two CCUs.
This game theoretic model is underpinned by a queueing model that takes into account the stochastic nature of queueing systems.
This work extends the application of game theoretic models already present in the literature to healthcare (\cite{li2002cooperative, xie2006note}).

@@

\begin{itemize}
    \item The assumptions as to the strategy space of our players is
        restrictive: a single threshold policy might not be optimal (although it
        is present in various pieces of literature on optimal control of
        queueing systems: \cite{naor1969regulation, shone2013comparisons}).
        Indeed, in reality critical care managers could have far more complex
        boundaries for their heuristic decision making;
    \item This model only assumes the presence of two players however in reality
        the system has a variety of stakeholders. Multiplayer systems could be
        worth considering. This would reflect health boards/hospitals in a
        concentrated area so that interactions are not just between two
        hospitals but between many.
    \item The restriction to pure strategies is influenced by discussions with
        ABUHB and also does not detract from the results presented thanks to the
        Theorem and Lemma of Section \ref{queueingandgamemodels}.  However,
        allowing for mixed strategies could also be of interest, corresponding
        to decision making that is not constant over time. Managers could
        alternate between a variety of behaviours: at time accepting patient
        despite being busy and at other times not;
    \item Patient length of stay is assumed to be dependent on the CCU at which
        they receive service. A further extension of the work would be to use
        the service rate from original CCU (prior to diversion). This would
        require a Markov chain with a higher dimensional state space. In
        practice this corresponds to patient morbidity corresponding to the
        original CCU.
\end{itemize}

@@

The code used in this work can be found here: \url{https://gist.github.com/anonymous/81effc06eea70a9e4e2f}.
The graphics for this paper were obtained using software described in \cite{Hunter:2007, sage} a worksheet with the data and code used for the plots can be found here: \url{https://cloud.sagemath.com/projects/c293aefd-1fdf-4b9c-95f4-75bb77035e42/files/MeasuringThePriceOfAnarchyInCCUInteractions.sagews}.

\newpage
\bibliography{library}
